A Motion Blur Resilient Fiducial Marker For Quadcopter Imaging

### 2014-03-01 ECCV/paper.tex
1. Synced with Michael's version 2. Added training data figure, recognition rate chart, tables and more ...

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{ruler}
\usepackage{color}
\usepackage{url}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\usepackage{multirow}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\captionsetup[table]{belowskip=12pt,aboveskip=4pt}
\setlength{\floatsep}{5pt plus 2pt minus 2pt}
\setlength{\textfloatsep}{5pt plus 2pt minus 2pt}
\setlength{\intextsep}{5pt plus 2pt minus 2pt}

@@

\title{Design of Blur Invariant Fiducial for Low Cost Quadcopter} % Replace
% with your title

@@

\begin{abstract}
Fiducial markers are commonly used to track an object in an unknown environment
and finds use in various applications in Virtual Reality, Medical imaging,
Surveys, etc. The performance of popular fiducials is satisfactory when there is
little motion or no motion in the device obtaining the imagery. But when there
is continuous and swift motion, as in the case of low cost quadcopters,
performance of these fiducials degrade significantly due to motion blur.
Inspired from Circular Data Matrix\cite{NaimarkF02}, we have designed a fiducial
that may be thought of as a binary code. It contains concentric white rings of
equal widths on a black background with a blurred border. Our fiducial detection
algorithm is based on the fact that there is no blur in the direction
perpendicular to the direction of the motion. The algorithm finds this
direction in a simple but accurate way. Then, we find ``signature'' of our
fiducial along this direction and further classify it to find the actual code
embedded in the fiducial. We will show through experimental validation that our
fiducial works better than popular fiducials such as ARTag under signifacant motion blur.

\keywords{Fiducials, Tracking, Blur Invariant}
\end{abstract}

\section{Introduction}
Recently, use of Unmanned Aerial Vehicles(UAV) such as quadcopters is increased
in various tracking activities e.g. aerial surveillance, search and rescue
operations etc. Navigation of these vehicles is mainly based on the
onboard inertial sensors. But sensor based measurments are prone to errors. So,
visual cues captured through onboard camera can be used along with the sensor
based measurments.  But, natural features may not be available always and
quality cannot be guranteed in every scene. So, results are not satisfactory in
uncontrolled environment and hence use of such features is limited specifically
when we need very high precision.

@@

\section{Related Work}
Our problem may be classified broadly in two areas; fiducial markers and
tracking. So, we will briefly discuss work done in both areas.
\begin{figure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{intersense.jpg}
  Intersense\cite{NaimarkF02}  
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}  
 \centering 
  \includegraphics[width=\linewidth]{pattKanji.pdf}
  ARToolkit\cite{ARToolkit02}  
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{ARtag.jpg}
  ARTag\cite{Fiala05}
 \end{subfigure} 
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{pifiducial.png}
  Pi-Tag\cite{Pitag13}  
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{our_fiducial}  
  Our Fiducial  
 \end{subfigure}
 \caption{Various Fiducial Marker Designs}
 \label{fig:previous_work}   
\end{figure}

The ARToolkit \cite{ARToolkit02} \cite{kato-artoolkit} is well known toolkit in
Augmented Reality system, widely used to find pose of the object on which it is
placed.  Fiala et al. \cite{Fiala05} developed fiducial named, ARTag, bi-tonal
system containing 2002 planar markers, each consisting of a square border and
an interior region filled with a 6x6 grid of black or white cells. It proved to
be more efficient than \cite{ARToolkit02} in terms of marker recognition rate
as well as the number of different patterns which can be created.

Concentric Rings are reportedly used first by Gatrell et al.\cite{concentric}
for monocular pose estimation as well as object identification in space. Cho et al.
\cite{Cho:2001}\cite{Cho97fastcolor} have used multi color rings in wide area
tracking in large scale applications. In \cite{NaimarkF02}, Naimark et al.
have developed ``Circular Data Matrix Fiducial System'', for wide area tracking.

@@

The main hurdle in detection of current fiducials is, the detecton of features;
lines in \cite{ARToolkit02}, corners in \cite{Fiala05} while ellipses in
\cite{Cho:2001}, \cite{Cho97fastcolor}, \cite{runetag11} and \cite{Pitag13}.
Such features are very difficult to detect accurately under significant motion
blur. So, under blur, recognition rate of these fiducials is very low.

@@

@@

\begin{table}
\caption{Recognition rate of AR Tag at various blur scales}
\centering
\begin{tabularx}{0.75\textwidth}{c *{6}{|Y}|}
\cline{2-6}
& \multicolumn{5}{ c |}{Recognition Rate (in percent)} \\
\cline{2-6} & \multicolumn{5}{ c |}{Blur Scale} \\ \cline{2-6}
& 30 & 35 & 40 & 45 & 50 \\ \cline{1-6}
\multicolumn{1}{ | c | }{ARTAg} & 100 & 19 & 9 & 5 & 0 \\ \cline{1-6}
\cline{1-6}
\end{tabularx}
\label{tab:ARTag_blur}
\end{table}

@@

\begin{table}
\caption{Blur angle range}
\centering
\begin{tabularx}{0.75\textwidth}{|Y|Y|}
\cline{1-2}
Blur Scale & Blur Angle Range \\ \cline{1-2}
30 & 0 -- 90\\ \cline{1-2}
35 & 32 -- 58 \\ \cline{1-2}
40 & 41 -- 48 \\ \cline{1-2}
45 & 41 -- 48 \\ \cline{1-2}
50 & 44 -- 46 \\ \cline{1-2}
\end{tabularx}
\label{tab:artag_blurangle}
\end{table}

@@

\begin{figure}
\includegraphics[width=\linewidth]{overall_flow.pdf}
\caption{Overall Workflow}
\end{figure}

Our fiducial detection strategy is different from \cite{NaimarkF02}\cite{Pitag13} and
works under significant amount of blur. Our algorithm is based on the fact that there
is no blur in the direction perpendicular to the direction of the motion.
Overall flow of fiducial detection algorithm can be summarised as follows:

\begin{itemize}
  \item Apply Gabor Filter on input image
  \item Find Connected components in Gabor output
  \item Cluster the connected components in bounding box
  \item Detect code in the bounding box
  \begin{itemize}
    \item Run PCA on Gabor output in bounding box
    \item Find intensity profile along first principal component passing through
    centroid
    \item If there are less than four black to white transitions in the
    intensity profile, ignore the bouding box
    \item Otherwise, classify the detected code by the classifer trained on
    standards.
  \end{itemize}
\end{itemize}

\textbf{Gabor Filter}: A 2D Gabor filter is a Gaussian kernel function modulated
by a sinusoidal plane wave. It is used to find high gradient patches from the
image. In our case, it will detect blur invariant sections of the fiducials. We
have applied Gabor filter for eight different orientations ($\theta = 0, 45,
90, 135, 180, 225, 270, 315$) and then calculated L2 norm of all outputs.

@@

\textbf{Classification}: Each synthetic fiducial pattern is blurred along 36
orientations (0, 10, 20, \ldots , 350) and intensity profile along first
principal component from every output is taken as training data for that fiducial pattern.
Training data for patterns having same number of rings is grouped together;
e.g., in two bit binary coded fiducial, training data for pattern ``01'' and
``10'' will be grouped together. Now, depending on the number of detected
rings, test pattern is matched against corresponding training data using
K-Nearest Neighbor Classifier (K=5).

\begin{figure}
\centering
  \includegraphics[width=.22\linewidth]{concentric_01_5_blurred_180.jpg}  
  \includegraphics[width=.22\linewidth]{concentric_01_5_blurred_45.jpg}
  \includegraphics[width=.22\linewidth]{concentric_01_5_blurred_90.jpg}  
  \includegraphics[width=.22\linewidth]{concentric_01_5_blurred_135.jpg}
  \caption{Sample images used to create training data for the fiducial pattern
  with binary code ``01''}
\end{figure}

@@

\begin{figure}
\centering
  \includegraphics[width=.8\linewidth]{output_00.jpg}
  \caption{Sample output for the fiducial pattern with binary code ``00''}
  \label{fig:output0}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=.8\linewidth]{output_01.jpg}
  \caption{Sample output for the fiducial pattern with binary code ``01''}
  \label{fig:output1}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=.8\linewidth]{output_10.jpg}
  \caption{Sample output for the fiducial pattern with binary code ``10''}
  \label{fig:output2}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=.8\linewidth]{output_11.jpg}
  \caption{Sample output for the fiducial pattern with binary code ``11''}
  \label{fig:output3}
\end{figure}

@@

\subsection{Comparison}
We will compare our results with standard fiducials such as ARTag. Also, we will
compare our results with Blur driven tracker(BLUT)\cite{Wu:2011}.
\subsubsection{Comparison with ARTag}

@@

\begin{figure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{11.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{12.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{13.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{14.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{15.jpg}
\end{subfigure}
\caption{Output of BlUT on sample image sequence}
\label{fig:BLUT_output}
\end{figure}

>>>>

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{ruler}
\usepackage{color}
\usepackage{url}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\usepackage{multirow}
\usepackage{tabularx}
\setlength{\extrarowheight}{2pt}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\captionsetup[table]{belowskip=12pt,aboveskip=4pt}
\setlength{\floatsep}{7pt plus 2pt minus 2pt}
\setlength{\textfloatsep}{7pt plus 2pt minus 2pt}
\setlength{\intextsep}{7pt plus 2pt minus 2pt}

@@

\title{A Motion Blur Resilient Fiducial Marker For Quadcopter Imaging} % Replace
% with your title

@@

\begin{abstract}
This paper describes the design and evaluation of a binary fiducial marker
design for use with low-cost quadcopters.  Fiducial markers are commonly placed
into environments to provide a reliable and unique scene marker conducive for
tracking and recognition.   For quadcopter applications, fiducials are
typically used for evaluating planning algorithms by allowing ground truth
positions in the scene to be detected from the quadcopter's camera. Quadcopters,
however, are subject to quick and unstable motions that causes notable motion
blur that severely affects the detection rate of existing fiducial markers. 
This motivated us to design a fiducial that is robust to motion blur. Our design
uses a series of concentric circles with the observation is that the direction
perpendicular to the motion blur direction will be unaffected by the blur and
therefore still be recognized as a binary code.  We detail the design and
detection algorithm for our fiducial and show that our marker can significantly
outperform existing fiducials on scenes captured with a quadcopter.
\keywords{Fiducials, Tracking, Detection using Blur}
\end{abstract}

\section{Introduction}
Recently, use of Unmanned Aerial Vehicles(UAV) such as quadcopters is increased
in various tracking activities e.g. aerial surveillance, search and rescue
operations etc. Navigation of these vehicles is mainly based on the
onboard inertial sensors. But sensor based measurments are prone to errors. So,
visual cues captured through onboard camera can be used along with the sensor
based measurments.  But, natural features may not be available always and
quality cannot be guaranteed in every scene. So, results are not satisfactory in
uncontrolled environment and hence use of such features is limited specifically
when we need very high precision.

@@

\section{Related Work}
Our work is related to two areas: fiducial markers and
object tracking. We will briefly discuss work done in both areas.
\begin{figure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{intersense.jpg}
  Intersense\cite{NaimarkF02}
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
 \centering
  \includegraphics[width=\linewidth]{pattKanji.pdf}
  ARToolkit\cite{ARToolkit02}
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{ARtag.jpg}
  ARTag\cite{Fiala05}
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{pifiducial.png}
  Pi-Tag\cite{Pitag13}
 \end{subfigure}
 \begin{subfigure}[b]{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{our_fiducial}
  Our Fiducial
 \end{subfigure}
 \caption{Various Fiducial Marker Designs}
 \label{fig:previous_work}
\end{figure}

The ARToolkit \cite{ARToolkit02} \cite{kato-artoolkit} is well known toolkit in
Augmented Reality system, widely used to find pose of the object on which it is
placed.  Fiala et al. \cite{Fiala05} developed a fiducial termed, ARTag, which is a
bi-tonal system consisting of a square border and
an interior region filled with a 6x6 grid of black or white cells. It proved to
be more efficient than \cite{ARToolkit02} in terms of marker recognition rate
as well as the number of different patterns which can be created.

Concentric rings are reportedly used first by Gatrell et al.\cite{concentric}
for monocular pose estimation as well as object identification in space. Cho et al.
\cite{Cho:2001}\cite{Cho97fastcolor} have used multi color rings in wide area
tracking in large scale applications. In \cite{NaimarkF02}, Naimark et al.
have developed ``Circular Data Matrix Fiducial System'', for wide area tracking.

@@

The main hurdle in detection of current fiducials is, the detecton of features:
lines in \cite{ARToolkit02}, corners in \cite{Fiala05} while ellipses in
\cite{Cho:2001}, \cite{Cho97fastcolor}, \cite{runetag11} and \cite{Pitag13}.
Such features are very difficult to detect accurately under significant motion
blur. So, under blur, recognition rate of these fiducials is very low.

@@

Before we detail the design of our fiducial, we examine the performance of two
popular fidicuals, ARTags and PiTags, in the face of motion blur.

@@

\begin{minipage}{\textwidth}
  \begin{minipage}[b]{0.49\textwidth}
\captionof{table}{Recognition rate of AR Tag at various blur scales}
\centering
 \begin{tabularx}{0.75\textwidth}{|Y|Y|}
\cline{1-2}
Blur Scale & Recognition Rate(\%) \\ \cline{1-2}
30 & 100\\ \cline{1-2}
35 & 19 \\ \cline{1-2}
40 & 9 \\ \cline{1-2}
45 & 5 \\ \cline{1-2}
50 & 0 \\ \cline{1-2}
\end{tabularx}
\label{tab:ARTag_blur}
 \end{minipage}	
 \begin{minipage}[b]{0.49\textwidth} 	
\captionof{table}{Blur angle range}
\centering
\begin{tabularx}{0.75\textwidth}{|Y|Y|}
\cline{1-2}
Blur Scale & Blur Angle Range \\ \cline{1-2}
30 & 0 -- 90\\ \cline{1-2}
35 & 32 -- 58 \\ \cline{1-2}
40 & 41 -- 48 \\ \cline{1-2}
45 & 41 -- 48 \\ \cline{1-2}
50 & 44 -- 46 \\ \cline{1-2}
\end{tabularx}
\label{tab:artag_blurangle}
 \end{minipage}
\end{minipage}

@@

@@

\begin{figure}
\includegraphics[width=\linewidth]{overall_flow.pdf}
\caption{Overall Workflow}
\end{figure}
Our fiducial detection strategy is different from \cite{NaimarkF02,Pitag13} and works
under significant amount of blur.   In particular, we assume that scene motion 
blur be modelled locally as a linear motion blur~\cite{}.  Under this assumption,
scene content perpendicular to the blur direction is unaffected by the blur.  Because
of our circular design, this means the perpendicular direction can be observed as a 
unblurred linear pattern.  Figure X shows an example with various motion directions.

Overall flow of fiducial detection algorithm can be summarised as follows:

\begin{itemize}
  \item Apply Gabor Filter on input image
  \item Find connected components in Gabor output
  \item Cluster the connected components into bounding boxes
  \item Detect code in the bounding box
  \begin{itemize}
    \item Run PCA on Gabor output in bounding box
    \item Find intensity profile along first principal component passing through
    centroid
    \item Classify the detected code by training examples
    standards.
  \end{itemize}
\end{itemize}

\textbf{Gabor Filter}: A 2D Gabor filter is a Gaussian kernel function modulated
by a sinusoidal plane wave. It is used to find high gradient patches from the
image. In our case, it will detect blur invariant sections of the fiducials. We
have applied Gabor filter for eight different orientations ($\theta = 0, 45,
90, 135, 180, 225, 270, 315$). We have used following parameters for
creating each Gabor kernel: $\lambda = 8$, $\gamma = 0.5$, $\sigma =
0.56\lambda$, $\psi = 0 \text(for real) = \pi/2 (for imaginary part)$. 
Then L2 norm of outputs along all orientations is calculated and finally, L2
normed image is binarized.

\begin{figure}
\centering
  \includegraphics[width=\linewidth]{training_data.pdf}  
  \caption{Training data for the fiducial pattern with binary code ``01''}
\end{figure}

@@

\textbf{Classification}: Each synthetic fiducial pattern is blurred along 36
orientations (0, 10, 20, \ldots , 350) and intensity profile along first
principal component from every output is taken as training data for that
fiducial pattern. Test pattern is matched against training data using 
K-Nearest Neighbor Classifier (K=5) to output class label.

To increase classification accuracy, training data for patterns having same
number of rings is grouped together; e.g., in two bit binary coded fiducial,
training data for pattern ``01'' and ``10'' will be grouped together, In three
bit binary coded fiducial, training data for pattern “001”, “010” and “100” will
form one group while training data for pattern “110”, “011” and “101” will be
in other group, etc. Now, Depending on the number of detected rings in test
pattern, it is matched against corresponding group of training data, again
using  K-Nearest Neighbor Classifier. In this way, if we detect either zero
rings or maximum possible rings in the test pattern, there will be no need to do
further classification.

@@

\begin{figure}
\begin{subfigure}{0.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{output_00.jpg}
  \caption{Binary code ``00''}
  \label{fig:output0}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{output_01.jpg}
  \caption{Binary code ``01''}
  \label{fig:output1}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{output_10.jpg}
  \caption{Binary code ``10''}
  \label{fig:output2}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{output_11.jpg}
  \caption{Binary code ``11''}
  \label{fig:output3}
  \end{subfigure}
  \caption{Sample Outputs for two bit binary coded fiducials}
\end{figure}

@@

\subsection{Comparison}
We will compare our results with standard fiducials such as ARTag. Also, we will
compare our results with Blur driven tracker(BLUT)\cite{Wu:2011}.
\subsubsection{Comparison with ARTag}
First, we have repeated blur simulation experiment on our fiducials too, i.e.,
we have downsized our fiducial to size 150x150 and blurred it along various
orientations with different blur scales. Then, we tried to detect the patterns
using algorithm presented in earlier section. The comparison of recongition
rate is shown in Fig. \label{fig:recognition_rate}.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{recognition_rate.png}
\caption{Comparison of recognition rate of AR Tag and our fiducials on
blur simulated data at various blur scales}
\label{fig:recognition_rate}
\end{figure}

Later, we recorded the real data using AR Drone quadcopter. In our experimental
setup, we have kept our fiducial alog with two AR Tags, to compare the
resilience of blur by each fiducial type. We have used ar\_track\_alvar, ROS
Wrapper for ALVAR libray \cite{ros_alvar}, to detect AR Tags from the stream
captured through quadcopter. The comparison of recognition rate shown in Table
\ref{tab:recongition_accuracy}. Classification accuracy of all fiducials(ARTag
as well as ours) is almost 100\%.

\begin{table}
\caption{Comparison of Recognition Rate of AR Tag and our fiducials on real
data captured through AR Drone}
\centering
\begin{tabularx}{0.8\textwidth}{|Y|Y|Y|Y|Y|}
\cline{1-5}
\multirow{2}{*}{AR Tag} & \multicolumn{4}{ c |}{Our Fiducial} \\ \cline{2-5}
& 00 & 01 & 10 & 11 \\  \cline{1-5}
65.6\% & 86.5\% & 94.1\% & 92.74\% & 93.54\% \\ \cline{1-5}
\end{tabularx}
\label{tab:recongition_accuracy}
\end{table}

@@

\begin{figure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{11.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{12.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{13.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{14.jpg}
\end{subfigure}
\begin{subfigure}[b]{.19\textwidth}
\includegraphics[width=\linewidth]{15.jpg}
\end{subfigure}
\caption{Output of BLUT on sample image sequence}
\label{fig:BLUT_output}
\end{figure}
